---
title: "CCE LTER Birds and Mammals"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Introduction to R and some of its uses.
---

```{r setup, include=FALSE}

# Load packages

library(learnr)
library(tidyverse)
library(plotly)
library(calecopal)

# Load data

calcofi_trans <- read_csv('CalCOFI_transect.csv')
calcofi_obs <- read_csv('CalCOFI_obs.csv')

nmfs_trans <- read_csv('NMFS_transect.csv')
nmfs_obs <- read_csv('NMFS_obs.csv')

cpr_trans <- read_csv('CPR_transect.csv')
cpr_obs <- read_csv('CPR_obs.csv')

species <- read_csv('species_codes.csv')
behavior <- read_csv('behavior_codes.csv')

# Wrangle data

calcofi <- inner_join(calcofi_trans, calcofi_obs)
nmfs <- inner_join(nmfs_trans, nmfs_obs)
cpr <- inner_join(cpr_trans, cpr_obs)

dat <- rbind(calcofi,nmfs) %>% 
  rbind(., cpr) %>% 
  left_join(species, by = c('Species'='code')) %>% 
  left_join(behavior, by = c('Behavior'='code')) %>% 
  mutate(date = lubridate::ymd(Date),
         year = lubridate::year(date))%>% 
  select(Cruise, Transect.number, Area, year, date, Season,
         taxon, Species, species, Count, behavior) %>% 
  rename(cruise = Cruise, transect = Transect.number, area = Area, season = Season,
         species_code = Species, count = Count) %>% 
  filter(!is.na(species)) %>% 
  mutate(behavior = factor(behavior, levels=c('Feeding','On/inthewater',
                                              'Inflight','Bow-riding',
                                              'Shipattracted','Shipfollowing')))

# Tutorial options:

tutorial_options(exercise.timelimit = 10)

```

## Raw data

These data are from the [CCE LTER](http://dx.doi.org/10.6073/pasta/0994b34ec9ce56b82d7b00bac995975c), showing bird and mammal counts from cruise transects.

*Farallon Institute Advanced Ecosystem Research, CalCOFI - Scripps Institution of Oceanography, California Current Ecosystem LTER, and B. Sydeman. 2017. Bird and mammal observations aboard CalCOFI (1987-2015, ongoing), NMFS (1996-2015, ongoing) and CPR (2003-2006, completed) cruises. ver 2. Environmental Data Initiative.*

```{r data, exercise=TRUE}

# Display first 10 rows of data
head(dat, 10)

```

## Species list

### Mammals

```{r mammal-species, exercise=TRUE}

mammals <- dat %>% 
  filter(taxon=='mammal')

unique(mammals$species)

```

### Birds

```{r bird-species, exercise=TRUE}

birds <- dat %>% 
  filter(taxon=='bird')

unique(birds$species)

```

## Behavior

### Mammals

```{r mammal-behavior, exercise=TRUE, fig.width=8, fig.height=4}

# Mammals

mammal_behavior <- dat %>% 
  filter(taxon=='mammal',
         !is.na(behavior),
         (behavior!='Inflight'))

cols <- cal_palette('kelp1', n = 6, type = 'continuous')
names(cols) <- levels(behavior$behavior)

mammal_behavior %>% 
  group_by(year, behavior) %>% 
  summarise(total_count = sum(count)) %>%
  group_by(year) %>% 
  mutate(n = sum(total_count)) %>% 
  ggplot(aes(x = year, y = total_count, fill = behavior)) +
  geom_bar(position = 'fill', stat = 'identity') +
  scale_fill_manual(name = 'Behavior', values = cols) +
  geom_text(aes(x = year, y = 1.07, label = n),
            angle = 90) +
  theme_minimal() +
  theme(axis.text.y = element_blank()) +
  scale_x_continuous(breaks = c(1990,1995,2000,2005,2010,2015)) +
  scale_y_continuous(expand = c(0,0.09)) +
  labs(x = 'Year', y = 'Proportion of Observations',
       title = 'Mammal Behavior Observations (1987-2015)')

```

### Birds

```{r bird-behavior, exercise=TRUE, fig.width=8, fig.height=4}

# Birds

bird_behavior <- dat %>% 
  filter(taxon=='bird',
         !is.na(behavior),
         behavior!='Bow-riding')

cols <- cal_palette('kelp1', n = 6, type = 'continuous')
names(cols) <- levels(behavior$behavior)

bird_behavior %>% 
  group_by(year, behavior) %>% 
  summarise(total_count = sum(count)) %>%
  group_by(year) %>% 
  mutate(n = sum(total_count)) %>% 
  ggplot(aes(x = year, y = total_count, fill = behavior)) +
  geom_bar(position = 'fill', stat = 'identity') +
  scale_fill_manual(name = 'Behavior', values = cols) +
  geom_text(aes(x = year, y = 1.10, label = n),
            angle = 90) +
  theme_minimal() +
  theme(axis.text.y = element_blank()) +
  scale_x_continuous(breaks = c(1990,1995,2000,2005,2010,2015)) +
  scale_y_continuous(expand = c(0,0.13)) +
  labs(x = 'Year', y = 'Proportion of Observations',
       title = 'Bird Behavior Observations (1987-2015)')

```



```{r eval=FALSE, include=FALSE}

library(plotly)

channel <- read_csv('channel_islands.csv') %>% 
  mutate(date = lubridate::ymd(time),
         year = lubridate::year(date))



plot_ly(mtcars,
        x = ~wt, y = ~hp, z = ~qsec,
        color = ~am, colors = c('#BF382A', '#0C4B8E')) %>%
  add_markers()%>%
  layout(scene = list(xaxis = list(title = 'Weight'),
                     yaxis = list(title = 'Gross horsepower'),
                     zaxis = list(title = '1/4 mile time')))

cols <- cal_palette('kelp1', n = 17, type = 'continuous')

plot_ly(channel,
        x = ~date, y = ~depth, z = ~size,
        color = ~species_name,
        colors = cols) %>% 
  add_markers()

        , colors = colors,
             marker = list(symbol = 'circle', sizemode = 'diameter'), sizes = c(5, 150),
             text = ~paste('Country:', country, '<br>Life Expectancy:', lifeExp, '<br>GDP:', gdpPercap,
                           '<br>Pop.:', pop))
fig <- fig %>% layout(title = 'Life Expectancy v. Per Capita GDP, 2007',
         scene = list(xaxis = list(title = 'GDP per capita (2000 dollars)',
                      gridcolor = 'rgb(255, 255, 255)',
                      range = c(2.003297660701705, 5.191505530708712),
                      type = 'log',
                      zerolinewidth = 1,
                      ticklen = 5,
                      gridwidth = 2),
               yaxis = list(title = 'Life Expectancy (years)',
                      gridcolor = 'rgb(255, 255, 255)',
                      range = c(36.12621671352166, 91.72921793264332),
                      zerolinewidth = 1,
                      ticklen = 5,
                      gridwith = 2),
               zaxis = list(title = 'Population',
                            gridcolor = 'rgb(255, 255, 255)',
                            type = 'log',
                            zerolinewidth = 1,
                            ticklen = 5,
                            gridwith = 2)),
         paper_bgcolor = 'rgb(243, 243, 243)',
         plot_bgcolor = 'rgb(243, 243, 243)')

```


